---
title: "Econ 211C Assignment 1"
author: "Aj Fontana"
date: "April 24, 2017"
output: html_document
---
#Number 1
##a:
Is the process weakly stationary and if so calculate the autocovariances

$Y_t=(1+2.4L+0.8L^2)*e_t$

$Y_t=e_t+2.4*e_{t-1}+.8*e_{t-2}$

$E[Y_t]=E[e_t]+E[2.4*e_{t-1}]+E[0.8*e_{t-2}]$

Assuming a mean zero MA(2) process$E[Y_t]=\mu=0$ Autocovariances can be shown to be $E[Y_t*Y_{t-j}]$. Expanding will give the following expressions:

$\rightarrow \gamma_0=(1+\theta_{1}^{2}+\theta_{2}^{2})*\sigma{2}=(1+2.4^2+0.8^2)=7.4$

$\rightarrow \gamma_1=(1+\theta_{1}+\theta_{1}*\theta_{2})*\sigma{2}=(2.4+2.4*(0.8))=4.32$

$\rightarrow \gamma_2=\theta_2*\sigma^{2}=0.8$

##b/c
Show the process is not invertible and find an invertible representation for it
$Y_t=(1+2.4L+0.8L^2)*e_t$ can be rewritten as $(1+0.4L)*(1+2L)$ We can see this lies outside the unit circle.

The invertible representation can be found by replacing the root 2 ($\lambda$) with its reciprocal. Also, the autocovariance functions need to be multiplied by $\eta^2$ in this case $2^2$ 

$(1+04.L)*(1+\frac{1}{2}L)$

Expanding: $1+0.5L+0.4L+0.2L^2 = (1+0.9L+0.2L^2)$

$\gamma_0=(1+0.9^2+0.2^2)*4=7.4$

$\gamma_1=(0.9+0.9*0.2)*4=4.32$

$\gamma_2=(0.2)*4=0.8$

#Number 2

```{r, include=TRUE}
### Number 2
n<-24000
sig<-.5
eps<-rnorm(n,0,sig)
epsilonmatrix<-matrix(eps,24,1000)
mu=.61
theta=.95

y1<-mu+epsilonmatrix[2:24,]+theta*epsilonmatrix[1:23,]
y2<-mu+epsilonmatrix[2:24,]-theta*epsilonmatrix[1:23,]
par(mfrow=c(2,1))
matplot(y1,type="l",main="Time Path for Y1",col=rgb(0,0,0,.2))
matplot(y2,type="l",main="Time Path for Y2",col=rgb(0,0,0,.2))
```


#Number 3
##a
Is this ARMA process weakly stationary?

We only need to look at the AR portion because all MA(p) models are stationary.

$Y_t=1.3Y_{t-1}-0.4Y_{t-2}+e_t$

$\phi(L)=1.3L-0.4L^2$

$\phi(L)= 1-1.3z+.4z^2$

$=(1-.8z)(1-.5z)$

$\frac{1}{\phi(z)}=\frac{1}{(1-.8z)(1-.5z)}=\frac{-1}{(1-.8z)}+\frac{2}{(1-.5z)}$

which can be rewritten using infinite sums which eventually gives us...

$$(\sum_{j=0}^{\infty}[-(0.8)^j+2(0.5)^j]L^j)e_t$$

$$\sum_{j=0}^{\infty}[-(0.8)^j+2(.25)^j]e_{t-j}$$
Or we could have just simply checked to see if the roots are outside the unit circle by looking at the reciprocal of $\lambda_j$ which shows...

$\frac{1}{0.8}=1.25$ and $\frac{1}{0.5}=2$

##b

```{r,include=TRUE}
matrix1<-matrix(0,nrow=5,ncol=5)
matrix1[1,]<-c(0.7,0,0.1,0.5,0.2)
matrix1[2:5,1:4]<-diag(4)
eigen(matrix1)$value
```

Because one of the eigenvalues is outside of the unit circle this is not invertible.

##c
See Attachment



##d





##e

```{r,include=TRUE}
ARMA<-function(n,phi,theta){
  p<-length(phi)
  q<-length(theta)
  r<-max(p,q)
  es<-rnorm(n+r)
  Y<-rep(0,n+r)
  Y[1:r]<-rnorm(r)
  for(i in (r+1):(r+n)){
    Y[i]<-t(phi)%*%Y[(i-1):(i-p)]+t(theta)%*%es[(i-1):(i-q)]+es[i]
  }
}

```


##f

```{r,include=TRUE}
ARMA(10000,2,3)
```



